Technical Report

### 2018-06-22 fully-abstract-comp/technical-report.tex
Fix wrong order of arguments in seal.
Update and simplify a proof.

      % This case was never used (?) so it is not necessary.
      % \wedge \\
      % \quad \npair[n']{\src{\vsc_S'},\src{\vsc_S},\vsc_T',\vsc_T} \in \lrexj(W' \oplus W_o)
    \end{array}
    \right\}\cup\\
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma,\sigma_\baddr,\sigma_\eaddr}},\\ \seal{\sigma,\sigma_\baddr,\sigma_\eaddr} \endarray \right)} 
    \; \middle| \;
    \begin{array}{l}
      [\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\gsigrets \cup \gsigcloss) \tand\\
      \forall \sigma' \in [\sigma_\baddr,\sigma_\eaddr] \ldotp \exists r \in \dom(\pwheap) \ldotp \\
      \quad \pwheap(r) = (\pure,\_,H_\sigma) \tand H_\sigma \; \sigma' \nequal (\lrv \circ \xi)
    \end{array}
    \right\} \cup \\
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\;
    \begin{array}{l}
      \begin{array}{r l l}
        \perm = \noperm & \Rightarrow & \npair{(\linear,\baddr,\eaddr)} \in \lrp(W) \wedge \\
        \perm \in \readAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \stackReadCond{W} \wedge \\
        \perm \in \writeAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \stackWriteCond{W}
      \end{array}
    \end{array}
    \right\} \cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\lin),\baddr,\eaddr,\aaddr)},\\ ((\perm,\lin),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
    \begin{array}{l}
      \begin{array}{r l l }
        \perm = \noperm & \Rightarrow & \npair{(\lin,\baddr,\eaddr)} \in \lrp(W) \wedge \\
        \perm \in \readAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \readCond{\lin,W} \wedge\\
        \perm \in \writeAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \writeCond{\lin,W} \wedge\\
        % we are excluding rwx pointers.
        \perm \neq \rwx \wedge \\
        % \perm = \rwx &\Rightarrow& \array[t]{l}\npair{(\{\rwx,\rx\},\baddr,\eaddr)} \in \execCond{\lin,W} \wedge \\
        %                            \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W} \endarray\\
        \perm = \rx &\Rightarrow& \array[t]{l}\npair{[\baddr,\eaddr]} \in \execCond{W} \wedge\\
        \npair{[\baddr,\eaddr]} \in \xReadCond{W} \wedge \\
                                  \lin = \normal \\ \endarray
      \end{array}
    \end{array}
    \right\}
  \end{array}
\]

@@

\begin{lemma}
  \label{lem:codereg-Hsigma-not-lrv}
  If
  \begin{itemize}
  \item $\dom(\code) \subseteq \ta$
  \item $\sigrets \subseteq \gsigrets$
  \item $\sigcloss \subseteq \gsigcloss$
  \item $n > 0$
  \item $\sigrets,\sigcloss\vdash\var{code}$
  \item $\sigma \in (\sigrets \uplus \sigcloss)$
  \end{itemize}
  then
  \begin{equation*}
    \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}.H_\sigma~ \sigma \not\nequal (\lrv \circ \xi)
  \end{equation*}
\end{lemma}
\begin{proof}
  Consider the following two cases:
  \begin{itemize}
  \item $\sigma \in \sigrets$:

    % In this case, we can choose a $b,e$ and construct a $\hat{W}$ such that $\npair[0]{\left(\retptrd(\baddr,\eaddr),((\rw,\linear),\baddr,\eaddr,\baddr-1)\right)}$ is not in $\lrv(\xi(\hat{W}))$, but it is in $\codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}.H_\sigma~ \sigma~\hat{W}$.
    % It suffices to make choices for $b,e$ and $\hat{W}$ such that the following hold:
    % \begin{itemize}
    % \item $\sigrets \subseteq \gsigrets$: assumption.
    % \item $\dom(\code) \subseteq \ta$: assumption.
    % \item $\exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]}$ such that $\pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}, \aaddr'+\calllen)$, $\dom(\ms_S) = \dom(\ms_T) = [\baddr,\eaddr]$, $\decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}}$, $\code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets$.
    %   It is not difficult to construct a $\hat{W}$ with a suitable $\spatialo$ region to satisfy these conditions.
    % \end{itemize}
    %  DOMI: what if code has no call or no seal?
    It suffices to observe that $\npair[0]{(0,0)} \in \lrv(\xi(\hat{W}))$ but not $\npair[0]{(0,0)} \in \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}.H_\sigma~ \sigma~ \hat{W}$.

  \item $\sigma \in \sigcloss$:
    By definition of $\codereg{}$, it suffices to show an $\npair[0]{(\vsc,\vsc')} \in (\lrvg{\trusted}(W) \setminus \lrv(W))$ with $\exec{\vsc}$.
    If we take $\vsc=\vsc' = ((\rx,\normal),b,e,a)$ then it suffices to show that we can take $[b,e]$ such that $[b,e] \subseteq \ta$, $\npair{[b,e]} \in\xReadCond{W}$ but not $\npair{[b,e]}\in\readCond{\normal,W}$.
    Choosing an appropriate world, this follows by definition of $\readCond{}$ and $\xReadCond{}$ from Lemma~\ref{lem:codereg-H-not-subs-lrv}.
  \end{itemize}
\end{proof}

@@

    Take $\sigma \in [\sigma_b,\sigma_e]$, we need to prove that $\sigma \not\in (\gsigrets \cup \gsigcloss)$.
    Assume that $\sigma \in (\sigrets \cup \sigcloss)$ and we will prove a contradiction.

    From $\npair{(w,\_)} \in \lrv(W_w)$, we get a $r \in \dom(\pwheap[W_w])$ such that $\pwheap[W_w](r) = (\pure,\_,H_\sigma)$ and $H_\sigma~\sigma \nequal \lrv$.

    From $\memSat{\ms_S,\stk,\ms_\stk,\_}{W_M}$, it follows that $\npair{(\overline{\sigma},\_,\_)} \in \lrheap(\pwheap[W_M])(W_{\var{heap}})$ for some $\overline{\sigma}$, $W_{\var{heap}}$ with $W_M = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}}$.

    From this, it follows that there exists a $R_\var{seal} : \dom(\activeReg{\pwheap[W_{\var{heap}}]}) \fun \powerset{\Seal}$, $\biguplus_{r \in \dom(\activeReg{\pwheap[W_{\var{heap}}]})} R_\var{seal}(r)) \subseteq \overline{\sigma}$ and $\dom(\pwheap[W_{\var{heap}}](r).H_\sigma) = R_\var{seal}(r)$ for all $r$.

    From $\dom(\pwheap[W_{\var{heap}}](r).H_\sigma) = R_\var{seal}(r)$ and the definedness of all the $\oplus$ operations on worlds, it follows that also $\dom(\pwheap[W_w](r).H_\sigma) = R_\var{seal}(r)$ for all $r$.

    From the facts that $\pwheap[W_w](r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}$, $\pwheap[W_w].H_\sigma~\sigma \nequal \lrv$, the disjointness of all the $R_{\var{seal}}(r)$, it follows that $r = r_\code$.
    However, this contradicts Lemma~\ref{lem:codereg-Hsigma-not-lrv}.

@@

@@

  \item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\gsigrets\cup \gsigcloss)$ then $\vsc$ is reasonable for $n-1$ steps: By $\npair{(w,\_)} \in \lrv(W_w)$, we get some region $r \in \dom(\pwheap)$ and a set of return seals, a set of closure seals and a code memory: $\sigrets',\sigcloss',\mscode'$ such that $\pwheap(r) = (\pure,\_,H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode' \; \gc)$ and $\npair[n']{(\sealed{\sigma,\vsc},\_)} \in H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode' \; \gc \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}$, so in particular for $n-1$.

    It also follows that $\sigrets \uplus \sigcloss \# \sigrets' \uplus \sigcloss'$.

    This gives us three cases for $\vsc$.
    The two first cases $\vsc = \retptrd(\_,\_)$ and $\vsc = \retptrc(\_,\_,\_)$ are easily discharged as the reasonability definition puts no requirements on return pointers.
    For the final case, $\sigma \in \sigcloss$ and either
    \begin{itemize}
    \item $\dom(\code') \mathrel{\#} \ta \tand \npair[n-1]{(\vsc,\_)} \in \lrv (W)$; or
    \item $\dom(\code') \subseteq \ta \tand \sigcloss' \subseteq \gsigcloss \npair{(\vsc,\_)} \in \lrvtrusted(W)$
    \end{itemize}
    and in either case $\sigma \in \sigcloss'$
    In the first case, the result follows from the induction hypothesis.
    In the second case, we have a contradiction with $\sigma \not\in (\gsigrets\cup \gsigcloss)$.
  \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$ and $n > 0$, then $\ms_S(\aaddr)$ is reasonable up to $n - 1$ steps for all $\aaddr \in ([\baddr,\eaddr] \setminus \ta)$:

>>>>

      % This case was never used (?) so it is not necessary.
      % \wedge \\
      % \quad \npair[n']{\src{\vsc_S'},\src{\vsc_S},\vsc_T',\vsc_T} \in \lrexj(W' \oplus W_o)
    \end{array}
    \right\}\cup\\
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma_\baddr,\sigma_\eaddr,\sigma}},\\ \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \endarray \right)} 
    \; \middle| \;
    \begin{array}{l}
      [\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\gsigrets \cup \gsigcloss) \tand\\
      \forall \sigma' \in [\sigma_\baddr,\sigma_\eaddr] \ldotp \exists r \in \dom(\pwheap) \ldotp \\
      \quad \pwheap(r) = (\pure,\_,H_\sigma) \tand H_\sigma \; \sigma' \nequal (\lrv \circ \xi)
    \end{array}
    \right\} \cup \\
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\;
    \begin{array}{l}
      \begin{array}{r l l}
        \perm = \noperm & \Rightarrow & \npair{(\linear,\baddr,\eaddr)} \in \lrp(W) \wedge \\
        \perm \in \readAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \stackReadCond{W} \wedge \\
        \perm \in \writeAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \stackWriteCond{W}
      \end{array}
    \end{array}
    \right\} \cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\lin),\baddr,\eaddr,\aaddr)},\\ ((\perm,\lin),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
    \begin{array}{l}
      \begin{array}{r l l }
        \perm = \noperm & \Rightarrow & \npair{(\lin,\baddr,\eaddr)} \in \lrp(W) \wedge \\
        \perm \in \readAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \readCond{\lin,W} \wedge\\
        \perm \in \writeAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \writeCond{\lin,W} \wedge\\
        % we are excluding rwx pointers.
        \perm \neq \rwx \wedge \\
        % \perm = \rwx &\Rightarrow& \array[t]{l}\npair{(\{\rwx,\rx\},\baddr,\eaddr)} \in \execCond{\lin,W} \wedge \\
        %                            \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W} \endarray\\
        \perm = \rx &\Rightarrow& \array[t]{l}\npair{[\baddr,\eaddr]} \in \execCond{W} \wedge\\
        \npair{[\baddr,\eaddr]} \in \xReadCond{W} \wedge \\
                                  \lin = \normal \\ \endarray
      \end{array}
    \end{array}
    \right\}
  \end{array}
\]

@@

% \begin{lemma}
%   \label{lem:codereg-Hsigma-not-lrv}
%   If
%   \begin{itemize}
%   \item $\dom(\code) \subseteq \ta$
%   \item $\sigrets \subseteq \gsigrets$
%   \item $\sigcloss \subseteq \gsigcloss$
%   \item $n > 0$
%   \item $\sigrets,\sigcloss\vdash\var{code}$
%   \item $\sigma \in (\sigrets \uplus \sigcloss)$
%   \end{itemize}
%   then
%   \begin{equation*}
%     \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}.H_\sigma~ \sigma \not\nequal (\lrv \circ \xi)
%   \end{equation*}
% \end{lemma}
% \begin{proof}
%   Consider the following two cases:
%   \begin{itemize}
%   \item $\sigma \in \sigrets$:

%     % In this case, we can choose a $b,e$ and construct a $\hat{W}$ such that $\npair[0]{\left(\retptrd(\baddr,\eaddr),((\rw,\linear),\baddr,\eaddr,\baddr-1)\right)}$ is not in $\lrv(\xi(\hat{W}))$, but it is in $\codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}.H_\sigma~ \sigma~\hat{W}$.
%     % It suffices to make choices for $b,e$ and $\hat{W}$ such that the following hold:
%     % \begin{itemize}
%     % \item $\sigrets \subseteq \gsigrets$: assumption.
%     % \item $\dom(\code) \subseteq \ta$: assumption.
%     % \item $\exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]}$ such that $\pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}, \aaddr'+\calllen)$, $\dom(\ms_S) = \dom(\ms_T) = [\baddr,\eaddr]$, $\decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}}$, $\code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets$.
%     %   It is not difficult to construct a $\hat{W}$ with a suitable $\spatialo$ region to satisfy these conditions.
%     % \end{itemize}
%     %  DOMI: what if code has no call or no seal?
%     It suffices to observe that $\npair[0]{(0,0)} \in \lrv(\xi(\hat{W}))$ but not $\npair[0]{(0,0)} \in \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}.H_\sigma~ \sigma~ \hat{W}$.

%   \item $\sigma \in \sigcloss$:
%     By definition of $\codereg{}$, it suffices to show an $\npair[0]{(\vsc,\vsc')} \in (\lrvg{\trusted}(W) \setminus \lrv(W))$ with $\exec{\vsc}$.
%     If we take $\vsc=\vsc' = ((\rx,\normal),b,e,a)$ then it suffices to show that we can take $[b,e]$ such that $[b,e] \subseteq \ta$, $\npair{[b,e]} \in\xReadCond{W}$ but not $\npair{[b,e]}\in\readCond{\normal,W}$.
%     Choosing an appropriate world, this follows by definition of $\readCond{}$ and $\xReadCond{}$ from Lemma~\ref{lem:codereg-H-not-subs-lrv}.
%   \end{itemize}
% \end{proof}

@@

    This follows directly from $\npair{(w,\_)} \in \lrv(W_w)$ by definition of $\lrv$.
    % Take $\sigma \in [\sigma_b,\sigma_e]$, we need to prove that $\sigma \not\in (\gsigrets \cup \gsigcloss)$.
    % Assume that $\sigma \in (\sigrets \cup \sigcloss)$ and we will prove a contradiction.

    % From $\npair{(w,\_)} \in \lrv(W_w)$, we get a $r \in \dom(\pwheap[W_w])$ such that $\pwheap[W_w](r) = (\pure,\_,H_\sigma)$ and $H_\sigma~\sigma \nequal \lrv$.

    % From $\memSat{\ms_S,\stk,\ms_\stk,\_}{W_M}$, it follows that $\npair{(\overline{\sigma},\_,\_)} \in \lrheap(\pwheap[W_M])(W_{\var{heap}})$ for some $\overline{\sigma}$, $W_{\var{heap}}$ with $W_M = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}}$.

    % From this, it follows that there exists a $R_\var{seal} : \dom(\activeReg{\pwheap[W_{\var{heap}}]}) \fun \powerset{\Seal}$, $\biguplus_{r \in \dom(\activeReg{\pwheap[W_{\var{heap}}]})} R_\var{seal}(r)) \subseteq \overline{\sigma}$ and $\dom(\pwheap[W_{\var{heap}}](r).H_\sigma) = R_\var{seal}(r)$ for all $r$.

    % From $\dom(\pwheap[W_{\var{heap}}](r).H_\sigma) = R_\var{seal}(r)$ and the definedness of all the $\oplus$ operations on worlds, it follows that also $\dom(\pwheap[W_w](r).H_\sigma) = R_\var{seal}(r)$ for all $r$.

    % From the facts that $\pwheap[W_w](r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}$, $\pwheap[W_w].H_\sigma~\sigma \nequal \lrv$, the disjointness of all the $R_{\var{seal}}(r)$, it follows that $r = r_\code$.
    % However, this contradicts Lemma~\ref{lem:codereg-Hsigma-not-lrv}.

@@

    \dominique{TODO: we don't know that a in dom(code)!}

@@

  \item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\gsigrets\cup \gsigcloss)$ then $\vsc$ is reasonable for $n-1$ steps:

    By $\npair{(w,\_)} \in \lrv(W_w)$, we get some region $r \in \dom(\pwheap)$ and a set of return seals, a set of closure seals and a code memory: $\sigrets',\sigcloss',\mscode'$ such that $\pwheap(r) = (\pure,\_,H_\sigma)$ and $H_\sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode' \; \gc)$ and $\npair[n']{(\sealed{\sigma,\vsc},\_)} \in H_\sigma \; \sigma \; \xi^{-1}(W)$ for all $n' < n$, so in particular for $n' = n-1$.
    It follows easily from the above that also $\npair[n-1]{(\vsc,\_)} \in H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode' \; \gc \; \sigma \; \xi^{-1}(W)$.

    It also follows from $\memSat{\ms_S,\stk,\ms_\stk,\_}{W_M}$, $\sigma \not\in (\gsigrets\cup \gsigcloss)$ and $\sigrets \subseteq \gsigrets$ and $\sigcloss \subseteq \gsigcloss$ that $\sigrets \uplus \sigcloss \# \sigrets' \uplus \sigcloss'$.

    This gives us three cases for $\vsc$.
    The two first cases $\vsc = \retptrd(\_,\_)$ and $\vsc = \retptrc(\_,\_,\_)$ are easily discharged as the reasonability definition puts no requirements on return pointers.
    For the final case, $\sigma \in \sigcloss'$ and either
    \begin{itemize}
    \item $\dom(\code') \mathrel{\#} \ta \tand \npair[n-1]{(\vsc,\_)} \in \lrv (W)$; or
    \item $\dom(\code') \subseteq \ta$, $\sigcloss' \subseteq \gsigcloss$, and $\npair{(\vsc,\_)} \in \lrvg{\trust}(W)$ for $\trust = \untrusted$ iff $\nonExec{\vsc}$.
    \end{itemize}
    In the first case, the result follows from the induction hypothesis.
    In the second case, we have a contradiction with $\sigma \not\in (\gsigrets\cup \gsigcloss)$.

  \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$ and $n > 0$, then $\ms_S(\aaddr)$ is reasonable up to $n - 1$ steps for all $\aaddr \in ([\baddr,\eaddr] \setminus \ta)$:
